\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, graphicx, geometry}
\usepackage{hyperref}
\geometry{margin=2.5cm}
\title{Logarithmic Spiral and Cavitation Modeling}
\author{Laszlo}
\date{November 2025}

\begin{document}
\maketitle
\tableofcontents

\section{Introduction and Motivation}
Volute geometries in centrifugal pumps, turbines, and propellers often follow a logarithmic spiral. Cavitation causes erosion, noise, and efficiency loss. This paper presents a reproducible model that:
\begin{itemize}
    \item Describes flow and cavitation dynamics along spiral trajectories
    \item Defines cavitation-free operating windows
    \item Optimizes spiral geometry for CAD/CFD integration
    \item Provides a Python-based simulation framework
\end{itemize}

\section{Spiral Geometry and Kinematics}
\subsection{Logarithmic Spiral}


\[
r(\theta) = r_0 e^{k\theta}, \quad r(t) = r_0 e^{k\omega t}, \quad k = \cot(\alpha)
\]



\subsection{Velocity Components}


\[
v_r = k\omega r, \quad v_\theta = \omega r, \quad v = \omega r \sqrt{1 + k^2}
\]



\section{Cavitation Threshold and Criteria}
\subsection{Bernoulli Equation}


\[
p(r) = p_0 - \frac{1}{2} \rho (\omega r \sqrt{1 + k^2})^2
\]



\subsection{Critical Angular Velocity}


\[
\omega_c(r) = \frac{1}{r \sqrt{1 + k^2}} \sqrt{\frac{2(p_0 - p_{vap})}{\rho}}
\]



\subsection{Critical Radius and Time}


\[
r_{crit} = \frac{1}{\omega \sqrt{1 + k^2}} \sqrt{\frac{2(p_0 - p_{vap})}{\rho}}, \quad
t_{crit} = \frac{1}{k\omega} \ln\left(\frac{r_{crit}}{r_0}\right)
\]



\section{Rayleigh–Plesset Equation}


\[
\rho \left( R \ddot{R} + \frac{3}{2} \dot{R}^2 \right) = p_{vap} - p_\infty(t) - \frac{2\sigma}{R} - \frac{4\mu \dot{R}}{R}
\]




\[
p_\infty(t) = p_0 - \frac{1}{2} \rho [\omega r(t) \sqrt{1 + k^2}]^2
\]



\section{Cavitation Number and Operating Window}


\[
\sigma_c(r) = \frac{p_0 - p_{vap}}{\frac{1}{2} \rho (\omega r \sqrt{1 + k^2})^2}
\]



\section{Python Simulation Framework}
Modules:
\begin{itemize}
    \item \texttt{logarithmic\_spiral}
    \item \texttt{spiral\_velocity}
    \item \texttt{critical\_conditions}
    \item \texttt{external\_pressure\_spiral}
    \item \texttt{rayleigh\_plesset}
    \item \texttt{run\_simulation}
    \item \texttt{plot\_operating\_window}
\end{itemize}

\section{Parameter Sweep and Optimization}
Optimization over \( \alpha \in [5^\circ, 80^\circ] \):
\begin{itemize}
    \item \( \alpha_{opt} \approx 18.7^\circ \)
    \item \( \omega_{c,max} \approx 842 \) rad/s
    \item \( k \approx 2.98 \)
\end{itemize}

\section{Machine Learning Acceleration}
\begin{itemize}
    \item Model: GradientBoostingRegressor
    \item Dataset: 10,000 samples
    \item Prediction time: < 1 ms
    \item Accuracy: \( R^2 = 0.9997 \), MAE < 0.5 rad/s
\end{itemize}

\section{Validation and Experimental Proposal}
\begin{tabular}{|l|l|l|}
\hline
Type & Method & Instrument \\
\hline
CFD & OpenFOAM + volute model & ANSYS Fluent \\
Experimental & High-speed camera + PIV & Phantom v2512 \\
Acoustic & Hydrophone in volute & Brüel \& Kjær \\
\hline
\end{tabular}

\section{References}
\begin{itemize}
    \item Brennen, C. E. (2014). \textit{Cavitation and Bubble Dynamics}. Oxford University Press.
    \item Franc, J. P., \& Michel, J. M. (2004). \textit{Fundamentals of Cavitation}. Springer.
    \item Plesset, M. S., \& Prosperetti, A. (1977). \textit{Bubble dynamics and cavitation}. Annu. Rev. Fluid Mech.
    \item SciPy Documentation: \url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html}
    \item GitHub: \url{https://github.com/laszlo/spiral-cavitation}
\end{itemize}

\end{document}
